Does Covid encourage assaults?

### 2022-04-27 outputs/paper/paper.Rmd
most of structure is complete

---
title: "Modelling Covid's effect on Assaults in Toronto"
subtitle: "A statistical analysis of how the rise of Covid in 2020 affected assault rates in various downtown neighborhoods in Toronto"
author: 
  "Varun Vijay"
thanks: "Code and data are available at: https://github.com/Varun1005473462/final_folder-main.git."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper presents an analysis on the effect of Covid on assault rates in Toronto neighbourhoods in 2020. The paper uses datasets obtained from opendatatoronto to show the Covid's positive effect on the number assaults in a neighbourhood. The analysis consists of using the statistical programming language R to model the data using regression and constructing various plots and graphs to visually represent the data. THe results obtained will allow us to better predict people's behavior. With the better prediction of people's behaviour, the police can better manage the public order and government will be able to make reasonable and effective policy to reduce the crime rate.\\par\\textbf {Keywords:} Covid, Assaults, Crime,Toronto neighbourhoods, residents, open data toronto"

@@

The relevant datasets were obtained using opendatatoronto and show statistics related to Covid cases and to crime figures. The first dataset is about Covid cases in Toronto and records all Covid cases reported since January 2020. The dataset includes all cases related to outbreaks and sporadic cases, those that occur in the community. The data was extracted from the provincial Case & Contact Management System and the variables in the dataset include the neighbourhood the case was reported in, the date it was reported, the client gender, Age group and the classification of the case, whether it was probable or improbable. The full dataset /@ref(table:dataset1) can be found in Appendix B.

The second dataset used for this paper was the dataset related to crime statistics in Toronto neighbourhoods. The dataset contains a list of crime related statistics by neighbourhood and includes figures such as number of assaults, number of car thefts, number of shootings, etc. for the years 2014 to 2020. We are only interested in the number of assaults for 2019 and 2020. The full dataset /@ref(table:dataset2) can be found in Appendix B.

The datasets were extracted using R [@citeR] and R packages opendatatoronto [@citeopendatatoronto], tidyverse [@tidyverse], dpylr [@dpylr] and knitr [@knitr]. The data was then processed, manipulated and analysed for the purposes of this paper.

@@

The final dataset /@ref(table:dataset_final) consists of 140 observations and 13 variables:
Neighbourhood: Name of the neighbourhood
Assault_2019: number of assaults in that neighbourhood for the year 2019
Assault_2020: number of assaults in that neighbourhood for the year 2020
FEMALE_Covid_CASES: Number of reported covid cases in that neighbourhood for females
MALE_Covid_CASES: Number of reported covid cases in that neighbourhood for males
transgender_Covid_CASES: Number of reported covid cases in that neighbourhood for transgenders
unknown_covid_cases: Number of reported covid cases in that neighbourhood for people of unknown genders
other_covid_cases: Number of reported covid cases in that neighbourhood for people of other genders
<=19: number of reported covid cases for people less than 19 years old
20 to 29: number of reported covid cases for people between 20 and 29
30 to 39: number of reported covid cases for people between 30 and 39
40 to 49: number of reported covid cases for people between 40 and 49
Total_CASES: total number of covid cases in a neighbourhood

@@

## final model graphs

@@

```{r figure 5, echo=FALSE}
par(mfrow = c(2,2))
model<-
  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`+`40 to 49`,
     data = final_dataset)
plot(model)
```

@@

The results obtained from our model and analysis of the dataset seem to imply that the number of covid cases in a neighbourhood had a positive relationship with the number of assaults in that same neighbourhood in the year 2020. The final model we obtained is:
$$ y = 47.418	-0.858*x_1+1.059*x_2 +13.834*x_3 -1.964*x_4 $$
where

@@

>>>>

---
title: "Does Covid encourage assaults?"
subtitle: "A statistical analysis of how the rise of Covid in 2020 affected assault rates in various neighborhoods in Toronto"
author: 
  "Varun Vijay"
thanks: "Code and data are available at: https://github.com/Varun1005473462/final_folder-main.git."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper presents an analysis on the effect of Covid on assault rates in Toronto neighbourhoods in 2020. The paper uses datasets obtained from opendatatoronto to show the Covid's positive effect on the number assaults in a neighbourhood. The analysis consists of using the statistical programming language R to model the data using regression and constructing various plots and graphs to visually represent the data. THe results obtained will allow us to better predict people's behavior. With the better prediction of people's behaviour, the police can better manage the public order and government will be able to make reasonable and effective policy to reduce the crime rate.\\par\\textbf {Keywords:} Covid, Assaults, Crime,Toronto neighbourhoods, residents, open data toronto"

@@

The relevant datasets were obtained using opendatatoronto and show statistics related to Covid cases and to crime figures. The first dataset is about Covid cases in Toronto and records all Covid cases reported since January 2020. The dataset includes all cases related to outbreaks and sporadic cases, those that occur in the community. The data was extracted from the provincial Case & Contact Management System and the variables in the dataset include the neighbourhood the case was reported in, the date it was reported, the client gender, Age group and the classification of the case, whether it was probable or improbable. The full dataset \@ref(table:dataset1) can be found in Appendix B.

The second dataset used for this paper was the dataset related to crime statistics in Toronto neighbourhoods. The dataset contains a list of crime related statistics by neighbourhood and includes figures such as number of assaults, number of car thefts, number of shootings, etc. for the years 2014 to 2020. We are only interested in the number of assaults for 2019 and 2020. The full dataset \@ref(table:dataset2) can be found in Appendix B.

The datasets were extracted using R [@citeR] and R packages opendatatoronto [@citeopendatatoronto], tidyverse [@tidyverse], dplyr [@dplyr] and knitr [@knitr]. The data was then processed, manipulated and analysed for the purposes of this paper.

@@

The final dataset \@ref(table:dataset_final) consists of 140 observations and 13 variables:
Neighbourhood: Name of the neighbourhood
Assault_2019: number of assaults in that neighbourhood for the year 2019
Assault_2020: number of assaults in that neighbourhood for the year 2020
FEMALE_Covid_CASES: Number of reported covid cases in that neighbourhood for females
MALE_Covid_CASES: Number of reported covid cases in that neighbourhood for males
transgender_Covid_CASES: Number of reported covid cases in that neighbourhood for transgenders
unknown_covid_cases: Number of reported covid cases in that neighbourhood for people of unknown genders
other_covid_cases: Number of reported covid cases in that neighbourhood for people of other genders
<=19: number of reported covid cases for people less than 19 years old
20 to 29: number of reported covid cases for people between 20 and 29
30 to 39: number of reported covid cases for people between 30 and 39
40 to 49: number of reported covid cases for people between 40 and 49
Total_CASES: total number of covid cases in a neighbourhood

@@

## Final model graphs

@@

```{r figure 9, echo=FALSE}
par(mfrow = c(2,2))
model<-
  lm(Assault_2020 ~FEMALE_Covid_CASES+MALE_Covid_CASES+`30 to 39`+`40 to 49`,
     data = final_dataset)
plot(model)
```

@@

The results obtained from our model and analysis of the dataset seem to imply that the number of covid cases in a neighbourhood had a positive relationship with the number of assaults in that same neighbourhood in the year 2020. The final model we obtained is:
$$ y = 47.418	-0.858*x_1+1.059*x_2 +13.834*x_3 -1.964*x_4 $$
where 
y= number of assaults in a neighbourhood
x_1 = number of female covid cases in a neighbourhood
x_2 = number of male covid cases in a neighbourhood
x_3 = number of covid cases between ages 30 and 39
x_4 = number of covid cases between ages 40 and 49

@@

[@thereferencecanbewhatever]

[@ggplot2]

[@dplyr]

[@tidyr]

[@janitor]

[@reshape2]

[@here]

[@tinytex]

[@readr]

[@metaopendata]
